---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
```

```{css}
html, body {
  background-color: white;
  font-family: 'Calibri', 'Arial', sans-serif;
  color: black;
  margin: 20px;
  padding: 20px;
}

h1, h2, h3 {
  color: #574240;
  text-align: left;
}

h2 {
  font-size: 30px;
  margin-bottom: 10px;
}

h3 {
  font-size: 25px;
  margin-bottom: 10px;
}

p, li {
  font-size: 16px;
  line-height: 1.5;
}

```

## Where Did All My Money Go? - A Week in Review 💸
By Hani Patel

### About this report 
This dynamic report explores my personal spending over a week. It tracks every purchase I made, capturing:

- **Amount spent** – in dollars ($)
- **Category** – the type of purchase (Food, Transport, Entertainment, Clothes, Bills, etc.)
- **Type of spending** – whether it was a Need or a Want
- **Mood while spending** – how I felt at the time (Happy, Neutral, Bored, Stressed)

The purpose of this analysis is to:

* Identify which categories I spend the most on
* Compare spending between needs vs wants
* Explore whether mood might influence spending decisions

All data is self-reported, anonymous, and used only for this assignment.

```{r}
# Load tidyverse library quietly
suppressPackageStartupMessages(library(tidyverse)) 

# Load the data
# --------------------------------------------------------------------
logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRNu8aT_HxKQM7yf0QDIQjH6pJv-iOtRLI6-8Vln53WhwrSZsgylpXsS1scO-4gdaKZbuQIbIxjOwpw/pub?output=csv")

# --------------------------------------------------------------------
# Rename variables for clarity and create latest_data
# --------------------------------------------------------------------
latest_data <- logged_data %>% 
  rename(timestamp = Timestamp,
         amount = `How much did you spend ($)?`,
         category = `What did you spend your money on?`,
         type = `What type of spending was this?`,
         mood = `Mood while spending?`)

# --------------------------------------------------------------------
# Summary values
# --------------------------------------------------------------------
# Numeric summaries
total_spent <- sum(latest_data$amount) %>% round(2)
average_spent <- mean(latest_data$amount, na.rm = TRUE) %>% round(2)
highest_single_purchase <- max(latest_data$amount, na.rm = TRUE) %>% round(2)

# Categorical summaries
most_common_category <- latest_data %>% 
                        count(category, sort = TRUE) %>%
                        slice(1) %>%
                        pull(category)

most_common_mood <- latest_data %>% 
                    count(mood, sort = TRUE) %>%
                    slice(1) %>%
                    pull(mood)

# Spending by type (Need vs Wants)
needs_total <-sum(latest_data$amount[latest_data$type == "Need"]) %>% round(2)
wants_total <- sum(latest_data$amount[latest_data$type == "Want"]) %>% round(2)
```

### Key spending summary 📌

In total, I spent $`r total_spent` over the week.

On average, each purchase was $`r average_spent`, and my largest single purchase was $`r highest_single_purchase`.

The most common spending category was **`r most_common_category`**, and the mood I recorded most often was `r most_common_mood`.

I spent **$`r needs_total`** on needs and **$`r wants_total`** on wants this week. 

### Chart 1: How often did I spend in each category? 

```{r}
# Chart 1: Number of Purchases by Category 
category_plot <- latest_data %>% 
      ggplot() +
      geom_bar(aes(x = category), fill = "#FD727B") +
      labs(title = "Number of Purchases by Category", 
              x = "Category", 
              y = "Number of Purchases",
              caption = "Source: Self-reported spending data")

print(category_plot)
```

### Chart 2: Which categories cost me the most? 

```{r}
# Chart 2: Total spending per category
category_spending_plot <- latest_data %>% 
        group_by(category) %>% 
        summarise(total_spent = sum(amount)) %>%
        ggplot (aes(x = category, y = total_spent)) +
        geom_col(fill = "#FD72C1") +
        labs(title = "Total Spending per Category", 
             x = "Category", 
             y = "Total Amount Spent ($)",
             caption = "Source: Self-reported spending data")

print(category_spending_plot)
```

### Chart 3: What mood did I spend money in?

```{r}
# Chart 3: Number of Purchases by Mood
mood_plot <- latest_data %>% 
      ggplot() +
      geom_bar(aes(x = mood), fill = "#FDAE72") +
      labs(title = "Number of Purchases by Mood", 
             x = "Mood", 
             y = "Number of Purchases",
            caption = "Source: Self-reported spending data")

print(mood_plot)
```

### Final takeaway 💭 
Overall, this report suggests that my spending was mainly focused on **`r most_common_category`**, and I tended to spend most often when I felt **`r most_common_mood`**. This may suggest that my mood influences when or why I make purchases, although more data would be needed to confirm this.